Gaussian Processes Autoregressive Models for Forecasting the Disturbance Storm Time Index

### 2016-05-19 swsc.tex
Updates from ShareLaTeX

%Vision
In $Dst$ prediction a seemingly trivial yet highly informative prediction method known as the \emph{Persistence} model has been largely ignored in its crucial role in model evaluation. The \emph{Persistence} model is stated simply as using the previous value of $Dst$ as the prediction for the next time step ($\hat{Dst}(t) = Dst(t-1)$). Due to high auto correlation between $Dst(t)$ and $Dst(t-1)$ in the Omni data, the \emph{Persistence} model gives excellent predictive performance despite its apparent lack of sophistication. In essence it is a trivial hypothesis and hence must always be used as a base line to compare the performance of any proposed $Dst$ algorithm. Therefore in the context of OSA prediction, a model must outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.

%Assertion
It is with the above observation in mind, we propose \emph{Gaussian Process} models for OSA prediction of $Dst$. We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a litmus test.

% Talk about Gaussian Process
We use the hourly resolution Omni data set (see \cite{OmniPaper}) as extracted from NASA/GSFC's OMNI data set through OMNIWeb. The rest of the paper is organized as follows. Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and the various issues surrounding their application. In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $Dst$ model as well as other models outlined in \citet{Ji2012}.

>>>>

%Vision
In $Dst$ prediction a seemingly trivial yet highly informative prediction method is represented by the so called \emph{Persistence} model this is simply using the previous value of $Dst$ as the prediction for the next time step ($\hat{D}st(t) = Dst(t-1)$). Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), despite its apparent lack of sophistication the \emph{Persistence} model gives excellent predictive performance on error metrics. In essence it is a trivial hypothesis and hence it should always be used as a base line to compare the performance of any proposed $Dst$ algorithm. Moreover, given its zero computational cost, a model should at least outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.

%Assertion
In this paper, we propose \emph{Gaussian Process} models for OSA prediction of $Dst$. We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a base line.

% Talk about Gaussian Process
We use the hourly resolution data as extracted from NASA/GSFC's OMNI data set through OMNIWeb (see \cite{OmniPaper}). The rest of the paper is organized as follows. Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and some technical details on their application. In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $Dst$ model as well as the other models outlined in \citet{Ji2012}.
